Simultaneous Elicitation of Scoring Rule and Agent Preferences for Robust Winner Determination

### 2021-05-14 minimax.tex
header table 2
Some minor rephrasings

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We consider a set $A$ of $m$ alternatives (products, restaurants, public projects, job candidates, etc.), an infinite set $\N$ of potential agents.

@@

An important class of PSRs is the one using convex weights \cite{Stein1994,Llamazares2016}, meaning that the difference between the weight of the first position and the weight of the second position is at least as large as the difference between the weights of the second and third positions, etc.
\begin{equation} 
	\label{eq:convexity}
	\forall r \in \{1,\ldots,m-2\}: w_r - w_{r+1} \geq w_{r+1}-w_{r+2}.
\end{equation}
The constraint above is a natural and common assumption, often used when aggregating rankings in sport competitions (such as F1 racing, alpine skiing world cup); losing ranks at the top  is more damaging than losing ranks at the bottom.
Let $\W$ denote the set of such convex weight vectors.

Consider a specific finite set of agents $N^* \subset \N$, let $\profile^* = (\pref_j^*, j \in N^*)$ and $\w^*$ denote the profile and weight vector, unknown to us, that represent the preferences of the agents in $N^*$ and of the chair.

@@

Note that in \cref{claim:rankPMR}, in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\begin{proposition}\label{claim:PMR}
	The $\PMR$ can be written as:
	\begin{align} 
		\PMR(x,y; \pprofile, W)  
		& = \max_{w \in W} \sum_{j \in N} w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i 
	\end{align}
	where $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) has rank $r$ in the complete profile $\hat{\profile}$ defined in \cref{claim:rankPMR}. 
\end{proposition}
%\cref{claim:rankPMR} can be understood by observing that in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\cref{claim:PMR}  shows that PMR is linear in the weights.
%Recall that $\pw$ represents the weight vectors compatible with the preferences of the chair stated so far.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can thus be obtained by solving the following linear program defined on the variables $w_1, …, w_m$:
\begin{align}
	\max_{\w} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r} \quad
	\text{ s.t. } w_1 = 1 ≥ … ≥ w_m = 0, \text{\cref{eq:convexity}} \text{ and } \w \in \pw.
\end{align}
%Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
%(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).
The max regret $\MaxR(x; \pprofile, W)$ is determined by computing the pairwise regret of $x$ with all other alternatives in $A$, and the recommended alternatives are the ones with least max regret. 
Observe that when the $\PMR$ of an alternative $x$ (against some other alternative $y$) exceeds the best MR value found so far, we do not need to further evaluate $x$. 
This idea can be exploited using a minimax-search tree \cite{Braziunas2012}.

@@

\begin{example*}
	Suppose we want to ask the following question to the chair: $w_{2} - w_{3} ≥ 2 (w_{3} - w_{4})$. We show the profile in Figure \ref{fig:profileQstComm} to the chair and ask who should win (each column is a preference of an agent).	Both $a$ and $b$ have scores higher than $c$ and $d$ for all convex weights, thus either $a$ or $b$ will be picked under our hypothesis; and $s(a) ≥ s(b) ⇔ w_2 + 2 w_4 ≥ 3 w_3$.
	Figure \ref{fig:profileQstCommCompact} represents the same profile using a compressed view, the numbers in bold indicating the number of agents having the preference in the corresponding column. As the proof shows, constructed profiles require only four different linear orders.
	\begin{figure}
		\centering
		\begin{subfigure}[b]{0.49\textwidth}
			\begin{center}
				$
				\begin{array}{ccccccccc}
					c&d&d&a&a&a&b&b&b\\
					a&c&c&b&b&b&a&a&a\\
					b&b&b&c&c&c&d&d&d\\
					d&a&a&d&d&d&c&c&c\\
				\end{array}.
				$
			\end{center}
			\caption{}
			\label{fig:profileQstComm}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.49\textwidth}
			\begin{center}
				$
				\begin{array}{cccc}
					\mathbf{1}&\mathbf{2}&\mathbf{3}&\mathbf{3} \\
					c&d&a&b\\
					a&c&b&a\\
					b&b&c&d\\
					d&a&d&c\\
				\end{array}.
				$
			\end{center}
			\caption{}
			\label{fig:profileQstCommCompact}
		\end{subfigure}
		\caption{Profile representing a question to the chair in extended (a) and compact (b) form.}
	\end{figure}
\end{example*}

@@

\paragraph{Evaluation of Pessimistic Strategy}
\label{sec:lowRegret}
Our next set of experiments evaluate the Pessimistic strategy in absolute terms. 
We first wonder how many questions should be asked in order to achieve low regret, fixed at $n / 10$: this is equivalent to the difference of score of an alternative $x$ that results from switching from a profile $P$ to a profile $P'$ where a tenth of the agents rank $x$ last instead of first. Another interesting measure is the average number of questions asked to each participant before reaching zero regret. The results for various sizes are displayed in \cref{tab:questions}. The notation $q_{c {\scriptscriptstyle \MMR}≤ x}$ indicates the number of questions asked to the chair to reach a threshold regret $x$, averaged on $20$ runs. We use a similar notation for the questions to the voters $q_{v {\scriptscriptstyle \MMR}≤ x}$, but in this case we display the quartiles ([Q1|M|Q3]) of the number of questions asked to the voters, averaged on $20$ runs. It is important to mention that we stop our computation at $1000$ questions, the empty cells in \cref{tab:questions} indicate, thus, that the Pessimistic strategy applied to that dataset did not reach zero regret in $1000$ questions.
%For example, in the case $m = 10, n = 20$, Pessimistic asks on average $49$ questions to the chair and $18$ questions to each agent (\cref{tab:Numquestions} provides more data).
\vspace{-0.6cm}
\begin{table}
	\caption{Questions asked by Pessimistic strategy on several datasets. The column $q_{c \scriptscriptstyle{\MMR} ≤ n/10}$ ($q_{c \scriptscriptstyle{\MMR} = 0}$) indicates the avg number of questions asked to the chair to reach a threshold regret of $\frac{n}{10}$ (zero regret). The column $q_{v \scriptscriptstyle{\MMR} ≤ n/10}$ ($q_{v \scriptscriptstyle{\MMR} = 0}$) represents the avg of the quartiles of the questions asked to the voters to reach $\frac{n}{10}$ regret (zero regret).}
	\label{tab:questions}
	\begin{tabular}{cccS[table-number-alignment = right, table-figures-integer=2]@{ ± }S[table-number-alignment = left, table-figures-integer=2] S[table-figures-integer=3, table-figures-decimal=1]@{ | }S[table-figures-integer=2, table-figures-decimal=1]@{ | }S[table-figures-integer=2, table-figures-decimal=1]@{ ]} S[table-number-alignment = right, table-figures-integer=2]@{ ± }S[table-number-alignment = left, table-figures-integer=2]S[table-figures-integer=3, table-figures-decimal=1]@{ | }S[table-figures-integer=2, table-figures-decimal=1]@{ | }S[table-figures-integer=2, table-figures-decimal=1]@{ ]}}
		\toprule
		{dataset} & m & n &{$q_{c \scriptscriptstyle{\MMR} ≤ n/10}$} & {sd} & \multicolumn{3}{c}{$q_{v \scriptscriptstyle{\MMR} ≤ n/10}$} & {$q_{c \scriptscriptstyle{\MMR} = n/10}$} & {sd} & \multicolumn{3}{c}{$q_{v \scriptscriptstyle{\MMR} = 0}$} \\
		\midrule
		m5n20 & 5&20&0.0&0.0&[4.3 &4.95 & 5.84 &5.25&9.59&[ \ 5.36 & 6.15 & 7.21\\
		m10n20&10&20&0.0&0.0&[13.85 & 16.1& 18.41&31.95&26.62&[19.66 & 21.78 & 24.7\\
		m11n30&11&30&0.0&0.0&[16.55&19.0&22.26&45.15&40.14&[23.07&25.7&28.89\\
		tshirts&11&30&0.0&0.0&[12.92 & 16.17 &19.3\\
		courses&9&146&0.0&0.0&[6.35 &7.0 &7.0&0.0&0.0&[6.8 & 7.0 &7.0\\
		%m9n146&9&146&0&0&[1.94 &8 &9.25&999.3&0.47&0&0&[1.94&8&9.25&999.3&0.47\\
		m14n9&14&9&5.4&9.49&[30.3&33.45&36.65&64.05&40.37&[37.55&40.5&44.3\\
		skate&14&9&0.0&0.0&[11.35&11.8&12.25&0.0&0.0&[11.5&11.9&12.7 \\
		m15n30&15&30&0.0&0.0&[24.95&29.5&33.68&\multicolumn{2}{c}{---}&\multicolumn{3}{c}{---} \\

>>>>

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We consider a set $A$ of $m$ alternatives (products, restaurants, public projects, job candidates, etc.) and an infinite set $\N$ of potential agents.

@@

An important class of PSRs is the one using convex weights \cite{Stein1994,Llamazares2016}, meaning that the difference between the weight of the first position and the weight of the second position is at least as large as the difference between the weights of the second and third positions, etc.
\begin{equation} 
	\label{eq:convexity}
	\forall r \in \{1,\ldots,m-2\}: w_r - w_{r+1} \geq w_{r+1}-w_{r+2}.
\end{equation}
The constraint above is a natural and common assumption, often used when aggregating rankings in sport competitions (such as F1 racing, alpine skiing world cup): losing ranks at the top  is more damaging than losing ranks at the bottom.
Let $\W$ denote the set of such convex weight vectors.

We consider a specific finite set of agents $N^* \subset \N$ and let $\profile^* = (\pref_j^*, j \in N^*)$ and $\w^*$ denote the profile and weight vector, unknown to us, that represent the preferences of the agents in $N^*$ and of the chair.

@@

Note that in \cref{claim:rankPMR}, in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\begin{proposition}\label{claim:PMR}
	The $\PMR$ can be written as:
	\begin{align} 
		\PMR(x,y; \pprofile, W)  
		& = \max_{w \in W} \sum_{j \in N} w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i,
	\end{align}
	where $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) has rank $r$ in the complete profile $\hat{\profile}$ defined in \cref{claim:rankPMR}. 
\end{proposition}
%\cref{claim:rankPMR} can be understood by observing that in the case $(x \pprefeq_j y)$, $\beta$ is the number of alternatives incomparable with both $x$ and $y$.
\cref{claim:PMR}  shows that PMR is linear in the weights.
%Recall that $\pw$ represents the weight vectors compatible with the preferences of the chair stated so far.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can thus be obtained by solving the following linear program defined on the variables $w_1, …, w_m$:
\begin{align}
	\max_{\w} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r} \quad
	\text{ s.t. } w_1 = 1 ≥ … ≥ w_m = 0, \text{\cref{eq:convexity}} \text{ and } \w \in \pw.
\end{align}
%Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
%(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).
The max regret $\MaxR(x; \pprofile, W)$ is determined by computing the pairwise regret of $x$ with all other alternatives in $A$, and the recommended alternatives are the ones with least max regret. 
Observe that when the $\PMR$ of an alternative $x$ (against some other alternative $y$) exceeds the best MR value found so far, we do not need to further evaluate $x$. 
This idea can be exploited using a minimax-search tree \cite{Braziunas2012}.

@@

\begin{example*}
	Suppose we want to ask the following question to the chair: $w_{2} - w_{3} ≥ 2 (w_{3} - w_{4})$. We show the profile in Figure \ref{fig:profileQstComm} to the chair and ask who should win (each column is the preference of one agent).
	Both $a$ and $b$ have scores higher than $c$ and $d$ for all convex weights, thus either $a$ or $b$ will be picked under our hypothesis; and $s(a) ≥ s(b) ⇔ w_2 + 2 w_4 ≥ 3 w_3$.
	Figure \ref{fig:profileQstCommCompact} represents the same profile using a compressed view, the numbers in bold indicating the number of agents having the preference in the corresponding column. As the proof shows, constructed profiles require only four different linear orders.
	\begin{figure}
		\centering
		\begin{subfigure}[b]{0.49\textwidth}
			\begin{center}
				$
				\begin{array}{ccccccccc}
					c&d&d&a&a&a&b&b&b\\
					a&c&c&b&b&b&a&a&a\\
					b&b&b&c&c&c&d&d&d\\
					d&a&a&d&d&d&c&c&c\\
				\end{array}.
				$
			\end{center}
			\caption{}
			\label{fig:profileQstComm}
		\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.49\textwidth}
			\begin{center}
				$
				\begin{array}{cccc}
					\mathbf{1}&\mathbf{2}&\mathbf{3}&\mathbf{3} \\
					c&d&a&b\\
					a&c&b&a\\
					b&b&c&d\\
					d&a&d&c\\
				\end{array}.
				$
			\end{center}
			\caption{}
			\label{fig:profileQstCommCompact}
		\end{subfigure}
		\caption{Profile representing a question to the chair in extended (a) and compact (b) form.}
	\end{figure}
\end{example*}

@@

\paragraph{Evaluation of Pessimistic Strategy}
\label{sec:lowRegret}
Our next set of experiments evaluate the Pessimistic strategy in absolute terms. 
We first wonder how many questions should be asked in order to achieve low regret, fixed at $n / 10$: this is equivalent to the difference of score of an alternative $x$ that results from switching from a profile $P$ to a profile $P'$ where a tenth of the agents rank $x$ last instead of first.
Table 2, first five columns, contain the result: it displays, for each dataset, the number of questions asked to the committee, and the quartiles of the number of questions asked to the voters, averaged over $20$ runs. It is interesting to note that about twenty or thirty questions per voter on average suffice to reach a low regret in those instances. We find also noteworthy that the Pessimistic strategy chooses to ask zero questions to the chair but still achieves low regret, in most of those instances.

Another interesting measure is the average number of questions asked to each participant before reaching zero regret. The results for various sizes are displayed in the last two columns of \cref{tab:questions}.

 It is important to mention that we stop our computation at $1000$ questions, the empty cells in \cref{tab:questions} indicate, thus, that the Pessimistic strategy applied to that dataset did not reach zero regret in $1000$ questions.
\commentOC{Need to drop the explanation in Table 2 as this is now explained in the text; change qcMMR=n/10 to qcMMR=0, and add in the text the explanation about qxMMR=; change hyphens to blanks in the table,= I suppose.}
%For example, in the case $m = 10, n = 20$, Pessimistic asks on average $49$ questions to the chair and $18$ questions to each agent (\cref{tab:Numquestions} provides more data).
\vspace{-0.6cm}
\begin{table}
	\caption{Questions asked by Pessimistic strategy on several datasets. The column $q_{c \scriptscriptstyle{\MMR} ≤ n/10}$ ($q_{c \scriptscriptstyle{\MMR} = 0}$) indicates the avg number of questions asked to the chair to reach a threshold regret of $\frac{n}{10}$ (zero regret). The column $q_{v \scriptscriptstyle{\MMR} ≤ n/10}$ ($q_{v \scriptscriptstyle{\MMR} = 0}$) represents the avg of the quartiles of the questions asked to the voters to reach $\frac{n}{10}$ regret (zero regret).}
	\label{tab:questions}
	\begin{tabular}{cccS[table-number-alignment = right, table-figures-integer=2]@{ ± }S[table-number-alignment = left, table-figures-integer=2] S[table-figures-integer=3, table-figures-decimal=1]@{ | }S[table-figures-integer=2, table-figures-decimal=1]@{ | }S[table-figures-integer=2, table-figures-decimal=1]@{ ]} S[table-number-alignment = right, table-figures-integer=2]@{ ± }S[table-number-alignment = left, table-figures-integer=2]S[table-figures-integer=3, table-figures-decimal=1]@{ | }S[table-figures-integer=2, table-figures-decimal=1]@{ | }S[table-figures-integer=2, table-figures-decimal=1]@{ ]}}
		\toprule
		{dataset} & m & n &{$q_{c \scriptscriptstyle{\MMR} ≤ n/10}$} & {sd} & \multicolumn{3}{c}{$q_{v \scriptscriptstyle{\MMR} ≤ n/10}$} & {$q_{c \scriptscriptstyle{\MMR} = n/10}$} & {sd} & \multicolumn{3}{c}{$q_{v \scriptscriptstyle{\MMR} = 0}$} \\
		\midrule
		m5n20 & 5&20&0.0&0.0&[4.3 &4.95 & 5.84 &5.25&9.59&[ \ 5.36 & 6.15 & 7.21\\
		m10n20&10&20&0.0&0.0&[13.85 & 16.1& 18.41&31.95&26.62&[19.66 & 21.78 & 24.7\\
		m11n30&11&30&0.0&0.0&[16.55&19.0&22.26&45.15&40.14&[23.07&25.7&28.89\\
		tshirts&11&30&0.0&0.0&[12.92 & 16.17 &19.3\\
		courses&9&146&0.0&0.0&[6.35 &7.0 &7.0&0.0&0.0&[6.8 & 7.0 &7.0\\
		%m9n146&9&146&0&0&[1.94 &8 &9.25&999.3&0.47&0&0&[1.94&8&9.25&999.3&0.47\\
		m14n9&14&9&5.4&9.49&[30.3&33.45&36.65&64.05&40.37&[37.55&40.5&44.3\\
		skate&14&9&0.0&0.0&[11.35&11.8&12.25&0.0&0.0&[11.5&11.9&12.7 \\
		m15n30&15&30&0.0&0.0&[24.95&29.5&33.68&\multicolumn{2}{c}{---}&\multicolumn{3}{c}{---} \\
